\documentclass[12pt, letterpaper]{article}
\usepackage[utf8]{inputenc}

\usepackage{amsmath}
\title{Econ 520 Homework 1}
\author{David Zynda}
\date{August 29, 2018}


 
 \begin{document}
\maketitle
 \begin{enumerate}

	 \item  For each of the following experiments, describe the sample space. 
		 \begin{enumerate}
			 \item Toss a coin four times.\hfill \break 
				The sample space is all possible combinations of heads and tails for tossing one coin four times. For example, $$S = \{(H,H,H,H), (H,H,H,T), (H,H,T,H), ..., (T,T,T,T)\}$$ for a sample space consisting of 16 outcomes. 
			 \item Count the number of insect-damaged leaves on a plant. 
				\hfill \break
				The sample space is each possible number of insect damaged leaves. For example, $$S = \{1,2,...,n\}$$ for $n$ leaves on a plant. 
			 \item Measure the lifetime (in hours) of a particular brand of light bulb.
				\hfill \break 
				 This sample space could be anywhere from zero hours to, in theory, and infinite amount. Hence $S = \{0,1,2,...,\infty\}$
			 \item Record the weights of 10-day-old rats. 
				 \hfill \break
				 Let $\{x_n\}$ be the weight of the $n^{th}$ rat which will be in a range from $[0,\infty)$. Then, the sample space is $S = \{x_1, ... x_n\}$. 
			 \item Observe the proportion of defectives in a shipment of electronic components.
				 \hfill \break
				 Let $x_i = \frac{i}{n}$ where $n$ is the total number of components and $i$ represents the $i^{th}$ component from $1,..,n$. The, the sample space is $S = \{x_1, x_2, ... x_n\}$
		 \end{enumerate}
	 \item Suppose we toss two fair coins. Let $A$ be the event "two heads," let $B$ be the event "no heads," and let $C$ be the event "at least one head". Define the sample space and the probability function. Calculate the probabilities of the following events: \hfill \break
	Let $x$ denote the event that the flip is a head:
		 $P(x) = $
		 \[ \begin{cases}
			 \frac{1}{4} & x = 0 \\
			 \frac{1}{2} & x = 1 \\
			 \frac{1}{4} & x = 2 
		\end{cases}
		 \]

		\begin{enumerate}
			\item $A \cup B$: 
				$$S = \{(H,H),(T,T)\}$$
				$$P(A \cup B) = \frac{1}{4} + \frac{1}{4} = \frac{1}{2}$$
			\item $A \cap B $:
				$$S = \emptyset$$
				$$P(a \cap B) = 0$$
			\item $B \cap C$: 
				$$S = \emptyset$$
				$$P(B \cap C) = 0$$
			\item $A^c$: 
				$$S = \{(H,T), (T,H), (T,T)\}$$
				$$P = \frac{1}{4} + \frac{1}{4} + \frac{1}{4} = \frac{3}{4}$$
		\end{enumerate}
		
	\item For events A and B, find formulas for the proabilities of the following vents in terms of the quantities $P(A)$, $P(B)$, or $P(A \cap B)$:
		\begin{enumerate}
			\item either $A$ or $B$ or both:
			$$P(A) + P(B) - P(A \cap B)$$
			\item either $A$ or $B$ but not both:
			$$P(A) + P(B) - 2P(A \cap B)$$
			\item at least one of $A$ or $B$:
			$$P(A) + P(B) - P(A \cap B)$$
			\item at most one of $A$ or $B$:
			$$1 - p(A \cap B)$$
		\end{enumerate}		
	 \end{enumerate}



 \end{document}


